---
title: "ggplot2 demo"
subtitle: "Class 8"
author: "Maxwell Austensen"
date: "2018-10-28 (updated: `r Sys.Date()`)"
output:
  xaringan::moon_reader:
    lib_dir: libs
    nature:
      highlightStyle: github
      highlightLines: true
      countIncrementalSlides: false
---

```{r setup, include=FALSE}
options(htmltools.dir.version = FALSE, fig.height = 5, fig.width = 7)
library(tidyverse)
mtcars <- as_tibble(mtcars)
```


class: center, middle

# Plotting with `ggplot2` in **R**

---
Here is a sample dataset we'll be using for some examples.
```{r tidy=FALSE}
mtcars
```

---

```{r eval=FALSE, tidy=FALSE}
ggplot(data = mtcars) #<<
```

Every plot begins with the function `ggplot()` and providing the dataframe you want to use. 
---

```{r eval=FALSE, tidy=FALSE}
ggplot(data = mtcars) + 
  aes(x = mpg, y = disp) #<<
```

The second part of every plot is to specify the mapping of columns form the dataframe to aesthetic elements of the plot. This is done with the function `aes()`. Every function call for a ggplot2 plot is added together with `+`.

In the `aes()` function some of the most common aesthetic elements are:

* `x` for the x-axis
* `y` for the y-axis
* `color` for the color of data elements (lines, points, and outline of shapes)
* `fill` for the fill color of shapes
* `label` for data labels
* `group` for separate data elements by some grouping variable
* `shape` for the shape of data points
* `linetype` for the style
---

```{r eval=FALSE, tidy=FALSE}
ggplot(data = mtcars) + 
  aes(x = mpg, y = disp) + 
  geom_point() #<<
```

The third main element of every plot is the geometries, which determine how the data is displayed on the plot. The geometries are determined by the `geom_*()` functions. Some examples are:

* `geom_point()` for scatter plot
* `geom_line()` for line graphs
* `geom_bar()` for bar graphs, usually representing the number of rows by some grouping column
* `geom_col()` for bar graphs where the height of the bars is specified with a column of the dataframe
* `geom_histogram()` for histograms
* `geom_sf()` for a type of spatial dataframe called "simple features" for creating maps, which we'll cover later
* and many more!

---

```{r tidy=FALSE}
ggplot(data = mtcars) + 
  aes(x = mpg, y = disp) + 
  geom_point()
```

---
class: center, middle

# Some examples
---

```{r tidy=FALSE}
ggplot(mtcars) + 
  aes(x = mpg, y = wt) + #<<
  geom_point()
```

---

```{r tidy=FALSE}
ggplot(mtcars) + 
  aes(x = mpg, y = wt, color = cyl) + #<<
  geom_point()
```

---

```{r tidy=FALSE}
ggplot(mtcars) + 
  aes(x = mpg, y = wt, color = factor(cyl)) + #<<
  geom_point()
```

---

```{r tidy=FALSE}
ggplot(mtcars) + 
  aes(x = mpg, y = wt, color = factor(cyl), size = disp) + #<<
  geom_point()
```
---

```{r tidy=FALSE}
ggplot(mtcars) + 
  aes(x = mpg, y = wt, size = disp) + #<<
  geom_point()
```

---
```{r tidy=FALSE}
ggplot(mtcars) + 
  aes(x = mpg, y = wt) + 
  geom_point() #<<
```

---
```{r tidy=FALSE}
ggplot(mtcars) + 
  aes(x = cyl) + 
  geom_bar() #<<
```

---

```{r message=FALSE, tidy=FALSE}
ggplot(mtcars) + 
  aes(x = mpg) + 
  geom_histogram() #<<
```

---

```{r tidy=FALSE}
ggplot(mtcars) + 
  aes(y = mpg) + 
  geom_boxplot() #<<
```

---

You can apply titles and labels to you plot using `labs()`:

```{r fig.height=4, fig.width=8, tidy=FALSE}
ggplot(mtcars) + 
  aes(x = mpg, y = wt, color = factor(cyl)) +
  geom_point() +
  labs( #<<
    title = "Miles/Gallon by Weight of Car", #<<
    subtitle = "1973-74 models", #<<
    x = "Weight", #<<
    y = "Miles per Gallon", #<<
    color = "Number of cylinders", #<<
    caption = "Source: 1974 Motor Trend US magazine" #<<
  ) #<<
```

---
class: center, middle

# Live Demo!

